\chapter{Trusted Types integration into Create React App}

Create React App (CRA) (described in \ref{intro-cra}) was a second choice for integrating Trusted
Types. After taking a look in the implementation of CRA we found that this tool doesn't depend on
React directly, but only installs it after creating the project template. Our goal was to make sure
the generated project by CRA is Trusted Types compatible.

\section{Using Trusted Types compatible version of React}

To accomplish this, we would need to change the implementation of CRA to install Trusted Types
compatible version of React. Unfortunately, such version of React is implemented under a feature
flag which needs to be turned on at build time. The published version of React contains the already
built files with the feature flag turned off. This means that if you want to use the Trusted Types
compatible version of React you have to clone the React repository from sources, turn on the feature
flag and then build the framework. You can then use this version of React as a dependency in your
project.

Implementing this is non trivial, since it requires knowledge about React. More importantly though,
this is harder to maintain for the application authors since you need to keep up with the new
releases of React manually.

\section{Using Trusted Types compliant version of Webpack}

CRA internally uses webpack version 5 to provide fast development experience with convenient
features, such as hot reload. For production it handles minification and other optimizations. Some
of these features, especially hot reload breaks under Trusted Types in enforcement mode. Gladly,
webpack can be configured to be Trusted Types compliant by a small configuration change
\cite{webpack_tt_config}.

The problem with CRA is that the webpack configuration is abstracted away from the user. This means
that there is not a simple way how to change the configuration. One can find the webpack
configuration in the downloaded dependencies and change it there directly, but this won't really
work since those changes will be removed by the package manager when modifying the dependencies. In
general, one should never modify the contents of dependencies and treat it only as an output of a
package manager.

The optimal solution as of now is to use a package which allows you to spy on modules and change
what they return. This pattern is often used when mocking or spying in unit tests. However, it is
also a suitable workaround for this case \cite{cra_modify_webpack_config}. CRA team is reluctant to
change this, since the hidden configuration is a good default for the majority of users.

Another option is to use the \textit{eject} command from CRA. This is a one way operation and it
essentially unwraps all of the hidden configuration and creates these files in your project. You can
then simply edit the webpack configuration which is now part of your project. This is not
recommended and it should be used only when all other attempts fail. The generated webpack
configuration is pretty complex and updating it in the future might not be trivial.
