\chapter{Testing}


Trusted Types are agnostic to the testing framework used and type of tests. It is very common for
web application to have a combination of unit and end to end tests. The former are used to smaller
parts of application and typically run in nodeJS environment. The latter usually test the whole
application in browser environment and focus on the full application as a product.

\section{Cypress}

Cypress is one of the most popular frameworks for end to end testing of web applications
\cite{js_state:testing}. It enjoys a rich ecosystem of plugins and supporting software. It runs the
tests in real browsers and is fast and reliable due to its unique architecture.

Since the tests run in a browser developers can test the Trusted Types compliance of their
application. Cypress loads the user application in an iframe, in which Trusted Types can be enforced
without the need for Trusted Types compliance of Cypress itself. This means that Cypress commands
are allowed to use dangerous sinks, because they run outside of the iframe. This is extremely
important, because it makes all of the existing Cypress ecosystem work out of the box under Trusted
Types enforcement.

\subsection{Trusted Types plugin}

Even though Cypress supports Trusted Types out of the box, there are a few nuances the developer has
to overcome to be able to test the application and Trusted Types violations. For this reason, we
have created a Cypress plugin which abstracts the low level details and provides a nicer API for the
developers to use.

\begin{flushleft}\textbf {Removed CSP header}\end{flushleft}

Cypress removes the CSP header sent by the application server in order to load the application in an
iframe, which could be prevented by a \textit{scriptSrc} CSP directive. There is a workaround for
now, which is to intercept the request of the initial HTML application payload and copy the CSP
policy from the response headers to the response body inside a \textit{meta} tag.
\cite{cypress:csp_removal_issue}.

This is tedious for the developers, so we have created a custom command in our plugin which does
exactly this.

\bigskip
\begin{lstlisting}[language=JavaScript, caption=Intercept requests and enable CSP header inside via meta tag]
// NOTE: Based on https://glebbahmutov.com/blog/testing-csp-almost/
Cypress.Commands.add('enableCspThroughMetaTag', (options) => {
  const { urlPattern } = options ?? {};

  // Intercept all requests by default
  cy.intercept(urlPattern ?? '**/*', (req) => {
    return req.reply((res) => {
      const csp = res.headers['content-security-policy'];
      if (!csp || typeof res.body !== 'string') return;

      res.body = res.body
        .replace(
          new RegExp('<head>([\\s\\S]*)</head>'),
          new RegExp(`<head><meta http-equiv="Content-Security-Policy" content="${csp}">$1</head>`).toString()
        )
        // The following are needed because the regex replacement above inserts some characters
        .replace('/<head>', '<head>')
        .replace('<\\/head>/', '</head>');
    });
  }).as('enableCspThroughMetaTag');
});
\end{lstlisting}

\begin{flushleft}\textbf {Testing the violations}\end{flushleft}

While it is easy to test that the Trusted Types are supported (\ref{lst:cypress_test_tt_support}) it
is harder to test whether part of the code triggered Trusted Types violation. The reason is that the
violation does not produce any user visible behaviour because the DOM is unchanged and uncaught
exception is thrown and Cypress automatically fails the test. The developer needs to listen to
\textit{uncaught:exception} handler from Cypress and explicitely recognize the Trusted Types
violation based on the error thrown and tell Cypress that this error is intended so it doesn't fail
the test.

% To prevent the code to span two pages
\bigskip
\bigskip
\begin{lstlisting}[language=JavaScript, caption=Test Trusted Types support, label={lst:cypress_test_tt_support}]
it('supports TT', () => {
  expect(window.trustedTypes).not.to.be.undefined;
});
\end{lstlisting}

To simplify this low level handling for the developers we have created a custom command which
recognizes Trusted Types violations and remembers them for the lifetime of the current test so they
can be asserted later on.

\bigskip
\begin{lstlisting}[language=JavaScript, caption=Custom command to catch Trusted Types violations]
Cypress.Commands.add('catchTrustedTypesViolations', () => {
  if (catchTrustedTypesViolationsEnabled) return;
  catchTrustedTypesViolationsEnabled = true;
  cy.clearTrustedTypesViolations();

  // https://docs.cypress.io/api/events/catalog-of-events#To-catch-a-single-uncaught-exception
  cy.on('uncaught:exception', (err) => {
    const violationTypes = ['TrustedHTML', 'TrustedScript', 'TrustedScriptURL'] as const;
    const type = violationTypes.find((type) => err.message.includes(formatViolationMessage(type)));
    if (type) {
      trustedTypesViolations.push({
        message: formatViolationMessage(type),
        error: err,
      });
      // Return false to prevent the error from failing this test
      return false;
    }

    const policyCreationErrorMessage = formatViolationMessage('PolicyCreation');
    if (err.message.includes(policyCreationErrorMessage)) {
      trustedTypesViolations.push({
        message: policyCreationErrorMessage,
        error: err,
      });
      // Return false to prevent the error from failing this test
      return false;
    }
  });
});
\end{lstlisting}

The final part of the plugin is the API to assert the caught violations. There are multiple commands
which assert that certain types of violations happened. An example of a full Cypress test which
asserts the violation is listed in (\ref{lst:cypress_test}).

\bigskip
\begin{lstlisting}[language=JavaScript, caption=Example Trusted Types violation test, label={lst:cypress_test}]
it('assertTrustedTypesViolations', () => {
  cy.enableCspThroughMetaTag();
  cy.catchTrustedTypesViolations();
  cy.visit(url);

  cy.contains('unsafe html').click();
  cy.contains('unsafe html').click();
  cy.contains('duplicate policy').click();
  cy.contains('unsafe script').click();

  cy.assertTrustedTypesViolations(['TrustedHTML', 'TrustedHTML', 'PolicyCreation', 'TrustedScript']);
});
\end{lstlisting}

\begin{flushleft}\textbf {Releasing the plugin}\end{flushleft}

The plugin is implemented in a standalone repository \cite{cypress-trusted-types:github} and
published as an npm package \cite{cypress-trusted-types:npm} for everyone to use. Together with the
plugin API there is an example application and tests which showcase the plugin in action.
