\chapter{Testing frameworks integrations}


Trusted Types are agnostic to the testing framework used and the type of tests. It is very common
for web application to have a combination of unit and end to end tests. The former are used to
smaller parts of application and typically run in nodeJS environment. The latter usually test the
whole application in browser environment and test the application features from the user
perspective.

\section{Cypress Trusted Types plugin}

Even though Cypress (\ref{intro:cypress}) supports Trusted Types out of the box, there are a few
nuances the developer has to overcome to be able to test the application and Trusted Types
violations. For this reason, we have created a Cypress plugin which abstracts the low level details
and provides a nicer API for the developers to use.

\subsection{Removed CSP header}

Cypress removes the CSP header sent by the application server in order to load the application in an
iframe, which could otherwise be prevented by a \textit{scriptSrc} CSP directive. There is a
workaround for now, which is to intercept the request of the initial HTML application payload and
copy the CSP policy from the response headers to the response body inside a \textit{meta} tag.
\cite{cypress:csp_removal_issue}.

This is tedious for the developers, so we have created a custom command in our plugin which does
exactly this.

\bigskip
\begin{lstlisting}[language=JavaScript, caption=Intercept requests and enable CSP header inside via meta tag]
// NOTE: Based on https://glebbahmutov.com/blog/testing-csp-almost/
Cypress.Commands.add('enableCspThroughMetaTag', (options) => {
  const { urlPattern } = options ?? {};

  // Intercept all requests by default
  cy.intercept(urlPattern ?? '**/*', (req) => {
    return req.reply((res) => {
      const csp = res.headers['content-security-policy'];
      if (!csp || typeof res.body !== 'string') return;

      res.body = res.body
        .replace(
          new RegExp('<head>([\\s\\S]*)</head>'),
          new RegExp(`<head><meta http-equiv="Content-Security-Policy" content="${csp}">$1</head>`).toString()
        )
        // The following are needed because the regex replacement above inserts some characters
        .replace('/<head>', '<head>')
        .replace('<\\/head>/', '</head>');
    });
  }).as('enableCspThroughMetaTag');
});
\end{lstlisting}

\subsection{Testing the violations}

While it is easy to test that the Trusted Types are supported (\ref{lst:cypress_test_tt_support}) it
is harder to test whether some part of the code triggered Trusted Types violation. The reason is
that the violation does not produce any user visible behaviour because the DOM is unchanged and
uncaught exception is thrown and Cypress automatically fails the test. The developer needs to listen
to \textit{uncaught:exception} handler from Cypress and explicitely recognize the Trusted Types
violation based on the error thrown and tell Cypress that this error is intended so it doesn't fail
the test.

\bigskip
\begin{lstlisting}[language=JavaScript, caption=Test Trusted Types support, label={lst:cypress_test_tt_support}]
it('supports TT', () => {
  expect(window.trustedTypes).not.to.be.undefined;
});
\end{lstlisting}

To simplify this low level handling for the developers we have created a custom command which
recognizes Trusted Types violations and remembers them for the lifetime of the current test so they
can be asserted later on.

\bigskip
\begin{lstlisting}[language=JavaScript, caption=Custom command to catch Trusted Types violations]
Cypress.Commands.add('catchTrustedTypesViolations', () => {
  if (catchTrustedTypesViolationsEnabled) return;
  catchTrustedTypesViolationsEnabled = true;
  cy.clearTrustedTypesViolations();

  // https://docs.cypress.io/api/events/catalog-of-events#To-catch-a-single-uncaught-exception
  cy.on('uncaught:exception', (err) => {
    const type = violationTypes.find((type) => err.message.includes(quote(type)));
    if (type) {
      trustedTypesViolations.push({
        type,
        message: extractViolationMessage(err),
        error: err,
      });
      // Return false to prevent the error from failing this test
      return false;
    }
  });
});
\end{lstlisting}

The final part of the plugin is the API to assert the caught violations. There are multiple commands
which assert that certain types of violations happened. An example of a full Cypress test which
asserts the violation is listed in (\ref{lst:cypress_test}).

\bigskip
\begin{lstlisting}[language=JavaScript, caption=Example Trusted Types violation test, label={lst:cypress_test}]
it('assertTrustedTypesViolations', () => {
  cy.contains('unsafe html').click();
  cy.contains('unsafe html').click();
  cy.contains('duplicate policy').click();
  cy.contains('unsafe script').click();

  cy.assertTrustedTypesViolations([
    {
      type: 'TrustedHTML',
      message:
        "Failed to set the 'srcdoc' property on 'HTMLIFrameElement': This document requires 'TrustedHTML' assignment.",
    },
    {}, // No assertion is made for this violation
    {
      type: 'TrustedTypePolicyFactory',
      message: `Failed to execute 'createPolicy' on 'TrustedTypePolicyFactory': Policy with name "my-policy" already exists.`,
    },
    { type: 'TrustedScript' },
  ]);
});
\end{lstlisting}

\subsection{Releasing the plugin}

The plugin is implemented in a standalone repository \cite{cypress-trusted-types:github} and
published as an npm package \cite{cypress-trusted-types:npm} for anyone to use. Together with the
plugin API there is an example application and tests which showcase the plugin in action. The plugin
is to be added to the cypress community list of plugins for increased visibility.
